Parsing Postfix Log Files

### 2009-04-17 doc/logparser-results.tex
Corrections based on proofreading to the preface and chapters 1-3.

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  \Graphref{parsing time vs log file size vs number of log
lines graph} shows the parsing time in seconds, log file size in MB, and
number of log lines in tens of thousands, for each of the
\numberOFlogFILES{} log files.  \Graphref{parsing time vs log file size vs
number of log lines factor} plots both the ratio of log file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is more efficient in both cases); \tableref{parsing time vs log file size
vs number of log lines factor table} shows the same ratios for different
groups of log files.  The ratios are quite tightly banded, except they
increase (i.e.\ improve) for log files 22 \& 62--68, despite their larger
than usual size.  Both groups of log files are much larger than usual
because of mail loops caused by users who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loops have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the mean number of rules tried per
log line reduces, as does the mean parsing time per log line.
\Tableref{Number of rules per Postfix component} shows the number of rules
for each Postfix component; \graphref{Mean number of rules tried per log
line} shows the drop in the mean number of rules tried per log line for log
files containing a mail loop, and \graphref{Mean number of rules tried per
log line for each Postfix component} shows the mean number of rules tried
per log line for each Postfix component.

@@

\showgraph{build/graph-average-number-of-rules-tried-per-log-line}{Mean
number of rules tried per log line}{Mean number of rules tried per log
line}

\showgraph{build/graph-average-number-of-rules-tried-per-program}{Mean
number of rules tried per log line for each Postfix component}{Mean number
of rules tried per log line for each Postfix component}

@@

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time relative to shuffled ordering table} shows that
differences in rule ordering have less effect on parsing time when parsing
log files 22 \& 62--68, because of the different distribution of log lines
in those log files.  A careful examination of \graphref{Parsing time
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled ordering
graph}

\showtable{build/include-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled ordering
table}

@@

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules tried per log line is lower
(see \graphref{Mean number of rules tried per log line}).  The performance
of optimal ordering relative to the oracle is much worse when parsing the
first log file than for the remainder of the log files, because for the
first log file the hits field of every rule is zero, so optimal ordering
does not provide any benefit for that log file; the oracle, in contrast, is
flawless for every log file.

@@

\subsection{Scalability As The Number Of Rules Rises}

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  These mails did not produce warnings because they resemble
        child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.  XXX CHECK IF THERE IS
        ANY MENTION OF THIS IN THE COMPLICATION, AND ADD IF MISSING\@.

>>>>

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  \Graphref{parsing time vs log file size vs number of log
lines graph} shows the parsing time in seconds, log file size in MB, and
number of log lines in tens of thousands, for each of the
\numberOFlogFILES{} log files.  \Graphref{parsing time vs log file size vs
number of log lines factor} plots both the ratio of log file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is more efficient in both cases); \tableref{parsing time vs log file size
vs number of log lines factor table} shows the same ratios for different
groups of log files.  The ratios are quite tightly banded, except they
increase (i.e.\ improve) for log files 22 \& 62--68, despite their larger
than usual size.  Both groups of log files are much larger than usual
because of mail loops caused by users who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loops have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the mean number of rules tried per
log line reduces, as does the mean parsing time per log line.
\Tableref{Number of rules per Postfix component} shows the number of rules
for each Postfix component; \graphref{Mean number of rules used per log
line} shows the drop in the mean number of rules tried per log line for log
files containing a mail loop, and \graphref{Mean number of rules used per
log line for each Postfix component} shows the mean number of rules tried
per log line for each Postfix component.

@@

\showgraph{build/graph-average-number-of-rules-tried-per-log-line}{Mean
number of rules used per log line}{Mean number of rules used per log line}

\showgraph{build/graph-average-number-of-rules-tried-per-program}{Mean
number of rules used per log line for each Postfix component}{Mean number
of rules used per log line for each Postfix component}

@@

\Graphref{Parsing time of optimal and reverse orderings relative to
shuffled ordering graph} shows the parsing times of optimal and reverse
orderings relative to shuffled ordering; the mean relative parsing times
for different groupings of log files are given in \tableref{Parsing time of
optimal and reverse orderings relative to shuffled ordering table}.  This
optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table} shows that differences in rule ordering have less
effect on parsing time when parsing log files 22 \& 62--68, because of the
different distribution of log lines in those log files.  A careful
examination of \graphref{Parsing time of optimal and reverse orderings
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

\showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time of
optimal and reverse orderings relative to shuffled ordering}{Parsing time
of optimal and reverse orderings relative to shuffled ordering graph}

\showtable{build/include-optimal-and-reverse-vs-shuffle}{Parsing time of
optimal and reverse orderings relative to shuffled ordering}{Parsing time
of optimal and reverse orderings relative to shuffled ordering table}

@@

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules tried per log line is lower
(see \graphref{Mean number of rules used per log line}).  The performance
of optimal ordering relative to the oracle is much worse when parsing the
first log file than for the remainder of the log files, because for the
first log file the hits field of every rule is zero, so optimal ordering
does not provide any benefit for that log file; the oracle, in contrast, is
flawless for every log file.

@@

\subsection{Scalability As The Ruleset Grows}

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  These mails did not produce warnings because they resemble
        child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.

### 2009-04-17 doc/logparser.tex
Corrections based on proofreading to the preface and chapters 1-3.

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding Postfix log files, and subsequently in reconstructing the
    actions taken by Postfix when processing mail delivery attempts.  This
    thesis describes the creation of a parser for Postfix log files,
    documenting the architecture developed for this project and the parser
    that implements it, explaining the difficulties encountered and the
    solutions developed, with reference to an implementation that stores
    data gleaned from the log files in an SQL database.  The gathered data
    can be used to optimise current anti-spam measures, to produce
    statistics showing how effective those measures are, or to provide a
    baseline to test new anti-spam measures against.  The Postfix log file
    parser needs to be very precise and strict when parsing log files, yet
    must allow users to easily adapt or extend it to parse new log lines,
    without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  The newly developed
    architecture is designed to make the process of parsing new inputs as
    simple as possible, enabling users to trivially add new rules (to parse
    variants of known inputs) and relatively easily add new actions (to
    process a previously unknown category of inputs).  The parser
    implemented for this project is evaluated on the criteria of efficiency
    and coverage of Postfix log files, demonstrating that the conflicting
    goals of efficiency and accuracy can be balanced, and one need not be
    sacrificed to achieve the other.

    \SVN$Id$
    \begin{center}
        \SVNId{}
    \end{center}

>>>>

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding Postfix log files, and subsequently in reconstructing the
    actions taken by Postfix when processing mail delivery attempts.  This
    thesis describes the creation of a parser for Postfix log files,
    documenting the architecture developed for this project and the parser
    that implements it, explaining the difficulties encountered and the
    solutions developed, and evaluating the parser on the criteria of
    efficiency and Postfix log file coverage.  The parser stores data
    gleaned from the log files in an SQL database; the gathered data can be
    used to optimise current anti-spam measures, to produce statistics
    showing how effective those measures are, or to provide a baseline to
    test new anti-spam measures against.  The Postfix log file parser needs
    to be very precise and strict when parsing, yet must allow users to
    easily adapt or extend it to parse new log lines, without requiring
    that the user have an in-depth knowledge and understanding of the
    parser's internal workings.  The newly developed architecture is
    designed to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of known inputs) and relatively easily add new actions (to process a
    previously unknown category of inputs).  The parser implemented for
    this project is evaluated on the criteria of efficiency and coverage of
    Postfix log files, demonstrating that the conflicting goals of
    efficiency and accuracy can be balanced, and one need not be sacrificed
    to achieve the other.

    \SVN$Id$
